\vspec{} and Friends: A suite of utilities to model spectroscopic phase curves of 3D exoplanet atmospheres in the presence of stellar variability.

### 2024-07-30 src/tex/ms.tex
Add my own affiliations
Change some wording in examples

\author{Ted Johnson}
\affiliation{NASA Goddard Space Flight Center \\
8800 Greenbelt Rd \\
Greenbelt, MD 20771, USA}
\affiliation{Nevada Center for Astrophysics, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
% Also CRESST, SURA, SEEC!

@@

@@

In this section we will simulate the transit of a bare rock across a spotted star to demonstrate how an inhomogeneous
stellar surface can lead to a false atmospheric signal -- a phenomenon known as the ``transit light source effect'' \citep{rackham2018}.
During transit, the planetary atmosphere is illuminated by the portion of the star directly behind the planet. The depth of the transit
should be measured with respect for the spectrum of this region of the stellar surface. However, the disk-integrated stellar spectrum is often used instead.
In the case that the stellar surface is not homogeneous, the transit signal is contaminated by the star.
\citet{moran2023} first observed this effect using JWST from super-Earth GJ 486b.

@@

We will do three \vspec{} simulations:
\begin{enumerate}
    \item Bare rock, no spots.
    \item 1 bar \ce{H2O} atmosphere, no spots.
    \item Bare rock, 2.5\% coverage by 2700 K spots
\end{enumerate}

We expect in case 1 to observe a flat spectrum, and in case 2 to observe additional
absorption from the \ce{H2O} atmosphere. Note also that we disabled \ce{H2O}-\ce{H2O} collision-induced absorption (CIA) for
these simulations because the effect was not considered in the original analysis of GJ 486b.

@@

We find that, as expected, the addition of spots changes the transit depth in a way that depends on wavelength -- mimicking
light lost to absorbers in a planetary atmosphere. Figure \ref{fig:moran_transit} shows us that, in this case, we observe false absorption;
this is due to an increase in the spotted fraction of the surface relative to the quiet photosphere. However, if a spot was blocked by
the planet we might see similar features inverted to mimic emission.

\subsection{Phase Curve}
Analysis of JWST MIRI-LRS phase curves of GJ 1214b by \citet{kempton2023} did not consider stellar
contamination because of the slow rotation rate of the host star \citep[approximately 1/80$^{\text{th}}$ the orbital frequency,][]{cloutier2021}.
In this example we use \vspec{} to demonstrate that this is a reasonable assumption.

>>>>

\author{Ted Johnson}
\affiliation{Southeastern Universities Research Association, 1201 New York Avenue NW, Suite 430, Washington, DC 20055, USA}
\affiliation{Center for Research and Exploration in Space Science and Technology, NASA Goddard Space Flight Center, 8800 Greenbelt Road, Greenbelt, MD 20771, USA}
\affiliation{NASA Goddard Space Flight Center \\
8800 Greenbelt Rd \\
Greenbelt, MD 20771, USA}
\affiliation{Nevada Center for Astrophysics, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
\affiliation{Department of Physics and Astronomy, University of Nevada, Las Vegas, 4505 South Maryland Parkway, Las Vegas, NV 89154, USA}
% Also CRESST, SURA, SEEC!

@@

\author{Ashraf Dhahbi}
\affiliation{NASA Goddard Space Flight Center \\
8800 Greenbelt Rd \\
Greenbelt, MD 20771, USA}

\author{Tobi Hammond}
\affiliation{NASA Goddard Space Flight Center \\
8800 Greenbelt Rd \\
Greenbelt, MD 20771, USA}

@@

In this section we will simulate the transit of a bare rock across a spotted star to demonstrate how an inhomogeneous
stellar surface can lead to a false atmospheric signal -- a phenomenon known as the ``transit light source effect'' \citep{rackham2018}.
During transit, the planetary atmosphere is illuminated by the portion of the star directly behind the planet. The depth of the transit
should be measured with respect for the spectrum of this region of the stellar surface. However, the disk-integrated stellar spectrum is often used instead.
In the case that the stellar surface is not homogeneous, the transit signal is contaminated by the star.
We will reproduce results by \citet{moran2023}, who observed this effect with JWST/NIRSpec during a transit of super-Earth GJ 486b.

@@

We will do two \vspec{} simulations:
\begin{enumerate}
    \item 1 bar \ce{H2O} isothermal atmosphere, no spots. \label{case:1barh2o}
    \item No atmosphere, 2.5\% coverage by 2700 K spots \label{case:no_atmo}
\end{enumerate}
In Case \ref{case:1barh2o} we expect to see an absorption spectrum characteristic of \ce{H2O}. Case \ref{case:no_atmo} is more complicated, though in the absence of spots we would expect a flat spectrum. Note also that we disabled \ce{H2O}-\ce{H2O} collision-induced absorption (CIA) for
these simulations because the effect was not considered in the original analysis of GJ 486b.

@@

We find that, as expected, the addition of spots changes the transit depth in a way that depends on wavelength -- mimicking
light lost to absorbers in a planetary atmosphere. Figure \ref{fig:moran_transit} shows us that, in this case, we observe false absorption;
this is due to a temporary increase in the spotted fraction of the surface relative to the quiet photosphere. However, if a spot was blocked by
the planet we might see similar features inverted to mimic emission.

\subsection{Phase Curve}
Analysis of JWST MIRI-LRS phase curves of GJ 1214b by \citet{kempton2023} did not consider stellar
contamination because of the slow rotation rate of the host star \citep[approximately 1/80$^{\text{th}}$ the orbital frequency,][]{cloutier2021}.
In this example we simulate this system using \vspec{} and find that this is a reasonable assumption.
